MinCall --- MinION end2end convolutional deep learning basecaller

### 2017-07-01 paper/main.tex
Changed title
Minor fixes

\title{MinCall --- MinION end2end deep learning basecaller}
\author{Neven Miculinić \and Marko Ratković \and Mile Sikić\\
\texttt{\{neven.miculinic, marko.ratkovic, mile.sikic\}@fer.hr}}

@@

\section{Sequencing overview}
Conceptually, the MinION sequencer works as follows. It's a variation on now standard shotgun sequencing approach. First, DNA is sheared into smaller DNA fragments and adapters are ligated to either end of the fragments. The resulting DNA fragments pass through a protein embedded in a membrane via a nanometre-sized channel, a nanopore. A single DNA strand passes through the pore. Optionally, hairpin protein adapter can merge two DNA strands, allowing both template and complement read passing through the nanopore sequentially for more accurate reads. This technique is referred as 2D reads, while we focus on 1D reads containing only template DNA and no hairpin adapter.

@@

\section{Basecalling}
The core of the decoding process is the basecalling step, that is translating the current samples to the nucleotide sequence. Nowadays there are multiple basecalling options, what official and unofficial ones.

Earlier models were Hidden Markov model(HMM)-based where hidden state modeled DNA sequence of length 6 (6-mer) in the nanopore. Pore models were used in computing emission probabilities.~\cite{loman2015complete,schreiber2015analysis,szalay2015novo,timp2012dna} and the recent open source HMM-based basecaller Nanocall~\cite{david2016nanocall}. Modern basecallers use RNN base models, and we opted out using CNN instead with beam search.

We compared our model on R9 chemistry with Metrichor (HMM-based approach), Nanonet and DeepNano (RNN based approaches). For the detailed basecaller overview see the appendix~\ref{app:basecallers}.

@@

\emph{DeepNano}~\cite{Boza2017}  was the first open-source basecaller based on neural networks. It uses bidirectional recurrent neural networks implemented in Python, using the Theano library. When released, originally only supported R7 chemistry, but support for R9 and R9.4 was added recently.

>>>>

\title{MinCall --- MinION end2end convolutional deep learning basecaller}
\author{Neven Miculinić \and Marko Ratković \and Mile Sikić\\
\texttt{\{neven.miculinic, marko.ratkovic, mile.sikic\}@fer.hr}}

@@

\subsection{Sequencing overview}
Conceptually, the MinION sequencer works as follows. It's a variation on now standard shotgun sequencing approach. First, DNA is sheared into smaller DNA fragments and adapters are ligated to either end of the fragments. The resulting DNA fragments pass through a protein embedded in a membrane via a nanometre-sized channel, a nanopore. A single DNA strand passes through the pore. Optionally, hairpin protein adapter can merge two DNA strands, allowing both template and complement read passing through the nanopore sequentially for more accurate reads. This technique is referred as 2D reads, while we focus on 1D reads containing only template DNA and no hairpin adapter.

@@

\section{Related work}
The core of the decoding process is the basecalling step, that is translating the current samples to the nucleotide sequence. Nowadays there are multiple basecalling options, what official and unofficial ones.

Earlier models were Hidden Markov model (HMM)-based where hidden state modeled DNA sequence of length 6 (6-mer) in the nanopore. Pore models were used in computing emission probabilities.~\cite{loman2015complete,schreiber2015analysis,szalay2015novo,timp2012dna} and the recent open source HMM-based basecaller Nanocall~\cite{david2016nanocall}. Modern basecallers use RNN base models, and we opted out using CNN instead with beam search.

We compared our model on R9 chemistry with Metrichor (HMM-based approach), Nanonet\footnote{\url{https://github.com/nanoporetech/nanonet/}} and DeepNano~\cite{deepnano} (RNN based approaches). For the detailed basecaller overview see the appendix~\ref{app:basecallers}.

@@

\emph{DeepNano}~\cite{deepnano}  was the first open-source basecaller based on neural networks. It uses bidirectional recurrent neural networks implemented in Python, using the Theano library. When released, originally only supported R7 chemistry, but support for R9 and R9.4 was added recently.
